---
title: "Draft: Simulate DO data based on metabolism, then mess with that data to create case study examples"
author: "KAL"
date: "2025-04-17"

output:
   html_document:
    theme: simplex
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 13px;}
code.r{font-size: 9px;}
pre {font-size: 11px}
</style>
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/Publications/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```


```{r, echo = F, message = F, include=FALSE}
### Packages
lapply(c("dplyr","ggplot2", "lubridate","tidyverse", "streamMetabolizer", "patchwork","ggpubr"), require, character.only=T)
```


#### Again, this builds off the existing vinegette for metabolism data simulation

Available here: https://rdrr.io/github/USGS-R/streamMetabolizer/man/metab_sim.html


### Part 1: 
Simulate metabolism and sub-daily DO for the baseline 'true' metabolism regime, and modeled parameter estimates using the bayes model to model the 'true' metabolic state. 

#### STEP 1: Create the "true" daily metabolism values

```{r, echo = T, message = F, include=T}

set.seed(123)  # for reproducibility

dat_daily_T <- data.frame(
  date = seq(as.Date("2012-09-18"), as.Date("2012-09-28"), by = "1 day"),
  GPP.daily = round(rnorm(11, mean = 3, sd = 0.5),3), # some wobble
  ER.daily = round(rnorm(11, mean = -4, sd = 0.5),3),
  K600.daily = round(rnorm(11, mean = 21, sd = 0.01),2), # less wobble, could fix constant
  stringsAsFactors = FALSE
)

dat_daily_T

```

#### STEP 2: Get a 'true' sub-daily dataset structure

Start with metaboilizer's built in example: 

```{r, echo = T, message = F, include=T}
dat <- data_metab('10', res = '15') 
head(dat)
```

#### STEP 3: Simulate DO.obs using metab_sim

with some observation and process error

```{r, echo = T, message = F, include=T}
?metab_sim()

specs_sim <- specs(mm_name('sim'), err_obs_sigma=0.1, err_proc_sigma=2)

mm_sim <- metab_sim(
  specs = specs_sim,
  data = dat,
  data_daily = dat_daily_T
)

# Save simulated DO.obs
sim_do <- predict_DO(mm_sim)

head(sim_do)
```


#### STEP 4: Replace DO.obs in original data with simulated values

```{r, echo = T, message = F, include=T}
dat_simulated <- dat %>%
  select(-DO.obs) %>%
  left_join(sim_do[, c("solar.time", "DO.obs")], by = "solar.time")

head(dat_simulated)
```


#### STEP 5: Fit bayes model to the simulated data

```{r, echo = T, message = F, include=T}

specs_fit <- specs(mm_name('bayes', err_proc_iid = FALSE),
                   n_cores = 2, n_chains = 2, burnin_steps = 200, saved_steps = 100)

mm_fit <- metab(specs = specs_fit, data = dat_simulated)
params <- get_params(mm_fit)

```

##### Quick check model preformance

```{r, echo = F, message = F, include=F}
# extract fit:
T_mm_fit <- get_fit(mm_fit)
T_mm_fit
```

```{r, echo = T, message = F, include=T}
DOplot <-plot_DO_preds(predict_DO(mm_fit))
DOplot
```


```{r, echo = T, message = F, include=T}
metabplot<- plot_metab_preds(predict_metab(mm_fit))
metabplot
```
#### Convergence? 

```{r, echo = T, message = F, include=T}
get_fit(mm_fit)$overall %>%
  dplyr::select(ends_with('Rhat'))
```

#### STEP 6: Compare recovered values to the "true" ones

Aggregate sub daily do simulation with "true" metabolism regime into "true_df" 

```{r, echo = T, message = F, include=T}
true_df <- sim_do %>%
  left_join(dat_daily_T) %>%
  rename(GPP.true = GPP.daily, ER.true = ER.daily, K600.true = K600.daily)
```

Get model estimates of true metabolism parameters to see how well bayes model recovered intial estimates through RMSE

```{r, echo = T, message = F, include=T}
estimates <- params %>%
  rename(GPP.est = GPP.daily, ER.est = ER.daily, K600.est = K600.daily)

comp_df <- dat_daily_T %>%
  rename(GPP.true = GPP.daily, ER.true = ER.daily, K600.true = K600.daily)%>%
  left_join(estimates, by=c("date"))
```

#### Visualize:

How the modeled parameter estimates vary from the 'true' simulated metabolism regime to get baseline variation before creating other DO scenarios. 


```{r, echo = F, message = F, include=F, warning=F}
GPP_plot <- comp_df %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = GPP.true, color = "True")) +
  geom_point(aes(y = GPP.true, color = "True")) +

  # Estimated GPP
  geom_ribbon(aes(ymin = GPP.est - GPP.daily.sd,
                  ymax = GPP.est + GPP.daily.sd,
                  fill = "Modeled"),
              alpha = 0.2) +
  geom_line(aes(y = GPP.est, color = "Modeled")) +
  geom_point(aes(y = GPP.est, color = "Modeled")) +

  # Customize colors and labels
  scale_color_manual(name = "GPP Type",
                     values = c("True" = "darkgreen", "Modeled" = "goldenrod")) +
  scale_fill_manual(name = "GPP Type",
                    values = c("True" = "darkgreen", "Modeled" = "goldenrod")) +

  theme_bw() +
  ylab(expression(GPP~(g~O[2]~m^-2~d^-1)))

```

```{r, echo = T, message = F, include=T, warning=F}
fits <- comp_df%>% 
  group_by(date)%>%
  summarize(rmse = sqrt(mean((GPP.true-GPP.est)^2)),
              sd = sd(GPP.true),
              min = min(GPP.true, na.rm = T),
              max = max(GPP.true, na.rm = T),
              range = range(max-min, na.rm = T),
              nrmse = rmse/range)
```

```{r, echo = F, message = F, include=F, warning=F}

rmsehist_gpp <- ggplot(fits, aes(x = rmse)) +
  geom_histogram(aes(fill = ..x..), bins = 30) +
  scale_fill_gradient(low = "lightblue", high = "darkred") +
  theme_minimal() +
  labs(fill = "RMSE")

```




```{r, echo = F, message = F, include=F, warning=F}
ER_plot <- comp_df %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = ER.true, color = "True")) +
  geom_point(aes(y = ER.true, color = "True")) +

  # Estimated ER
  geom_ribbon(aes(ymin = ER.est - ER.daily.sd,
                  ymax = ER.est + ER.daily.sd,
                  fill = "Modeled"),
              alpha = 0.2) +
  geom_line(aes(y = ER.est, color = "Modeled")) +
  geom_point(aes(y = ER.est, color = "Modeled")) +

  # Customize colors and labels
  scale_color_manual(name = "ER Type",
                     values = c("True" = "purple4", "Modeled" = "goldenrod")) +
  scale_fill_manual(name = "ER Type",
                    values = c("True" = "purple4", "Modeled" = "goldenrod")) +

  theme_bw() +
  ylab(expression(ER~(g~O[2]~m^-2~d^-1)))

```

```{r, echo = F, message = F, include=T, warning=F}
fits <- comp_df%>% 
  group_by(date)%>%
  summarize(rmse = sqrt(mean((ER.true-ER.est)^2)),
              sd = sd(ER.true),
              min = min(ER.true, na.rm = T),
              max = max(ER.true, na.rm = T),
              range = range(max-min, na.rm = T),
              nrmse = rmse/range)
```

```{r, echo = F, message = F, include=F, warning=F}

rmsehist_ER <- ggplot(fits, aes(x = rmse)) +
  geom_histogram(aes(fill = ..x..), bins = 30) +
  scale_fill_gradient(low = "lightblue", high = "darkred") +
  theme_minimal() +
  labs(fill = "RMSE")

```


```{r, echo = F, message = F, include=F, warning=F}
K600_plot <- comp_df %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = K600.true, color = "True")) +
  geom_point(aes(y = K600.true, color = "True")) +

  # Estimated K600
  geom_ribbon(aes(ymin = K600.est - K600.daily.sd,
                  ymax = K600.est + K600.daily.sd,
                  fill = "Modeled"),
              alpha = 0.2) +
  geom_line(aes(y = K600.est, color = "Modeled")) +
  geom_point(aes(y = K600.est, color = "Modeled")) +

  # Customize colors and labels
  scale_color_manual(name = "K600 Type",
                     values = c("True" = "steelblue3", "Modeled" = "goldenrod")) +
  scale_fill_manual(name = "K600 Type",
                    values = c("True" = "steelblue3", "Modeled" = "goldenrod")) +

  theme_bw() +
  ylab(expression(K600~(g~O[2]~m^-2~d^-1)))

```

```{r, echo = F, message = F, include=T, warning=F}
fits <- comp_df%>% 
  group_by(date)%>%
  summarize(rmse = sqrt(mean((K600.true-K600.est)^2)),
              sd = sd(K600.true),
              min = min(K600.true, na.rm = T),
              max = max(K600.true, na.rm = T),
              range = range(max-min, na.rm = T),
              nrmse = rmse/range)
```

```{r, echo = F, message = F, include=F, warning=F}

rmsehist_K600 <- ggplot(fits, aes(x = rmse)) +
  geom_histogram(aes(fill = ..x..), bins = 30) +
  scale_fill_gradient(low = "lightblue", high = "darkred") +
  theme_minimal() +
  labs(fill = "RMSE")

```

```{r, echo = F, message = F, include=T}
ggarrange(
  GPP_plot,
  rmsehist_gpp, 
  ER_plot,
  rmsehist_ER, 
  K600_plot,
  rmsehist_K600,
  ncol = 2, nrow = 3,
  widths = c(0.7, 0.3),
  common.legend = T,
  align=c("v")
)
```





### Part 2: 
Wrong calibration example, same mini DOs one has static offset (-0.12), one doesn't the true simulated data is still the true data. 


#### STEP 1: 
 
```{r, echo = T, message = F, include=T}
dat_simulated_cal <- dat_simulated %>%
  mutate(DO.obs = DO.obs - 0.12)

head(dat_simulated_cal)
```



#### STEP 2: Fit bayes model to the calibration offset data

```{r, echo = T, message = F, include=T}
mm_fit_cal <- metab(specs = specs_fit, data = dat_simulated_cal)
params_cal <- get_params(mm_fit_cal)
```

##### Quick check model preformance

```{r, echo = F, message = F, include=F}
# extract fit:
Cal_mm_fit <- get_fit(mm_fit_cal)
Cal_mm_fit
```

```{r, echo = F, message = F, include=F}
DOplot <-plot_DO_preds(predict_DO(mm_fit_cal))
DOplot
```


```{r, echo = F, message = F, include=F}
metabplot<- plot_metab_preds(predict_metab(mm_fit_cal))
metabplot
```
#### Convergence? 

```{r, echo = F, message = F, include=F}
get_fit(mm_fit_cal)$overall %>%
  dplyr::select(ends_with('Rhat'))
```

#### STEP 3: Compare recovered values to the "true" ones

Aggregate sub daily do simulation with "true" metabolism regime into "true_df" 

```{r, echo = T, message = F, include=T}
true_df <- sim_do %>%
  left_join(dat_daily_T) %>%
  rename(GPP.true = GPP.daily, ER.true = ER.daily, K600.true = K600.daily)
```

Get model estimates of true metabolism parameters to see how well bayes model recovered intial estimates through RMSE

```{r, echo = T, message = F, include=T}
estimates <- params_cal %>%
  rename(GPP.est = GPP.daily, ER.est = ER.daily, K600.est = K600.daily)

comp_df <- dat_daily_T %>%
  rename(GPP.true = GPP.daily, ER.true = ER.daily, K600.true = K600.daily)%>%
  left_join(estimates, by=c("date"))
```

#### Visualize:

How the modeled parameter estimates vary from the 'true' simulated metabolism regime to get baseline variation before creating other DO scenarios. 


```{r, echo = F, message = F, include=F, warning=F}
GPP_plot <- comp_df %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = GPP.true, color = "True")) +
  geom_point(aes(y = GPP.true, color = "True")) +

  # Estimated GPP
  geom_ribbon(aes(ymin = GPP.est - GPP.daily.sd,
                  ymax = GPP.est + GPP.daily.sd,
                  fill = "Modeled"),
              alpha = 0.2) +
  geom_line(aes(y = GPP.est, color = "Modeled")) +
  geom_point(aes(y = GPP.est, color = "Modeled")) +

  # Customize colors and labels
  scale_color_manual(name = "GPP Type",
                     values = c("True" = "darkgreen", "Modeled" = "goldenrod")) +
  scale_fill_manual(name = "GPP Type",
                    values = c("True" = "darkgreen", "Modeled" = "goldenrod")) +

  theme_bw() +
  ylab(expression(GPP~(g~O[2]~m^-2~d^-1)))

```

```{r, echo = T, message = F, include=T, warning=F}
fits <- comp_df%>% 
  group_by(date)%>%
  summarize(rmse = sqrt(mean((GPP.true-GPP.est)^2)),
              sd = sd(GPP.true),
              min = min(GPP.true, na.rm = T),
              max = max(GPP.true, na.rm = T),
              range = range(max-min, na.rm = T),
              nrmse = rmse/range)
```

```{r, echo = F, message = F, include=F, warning=F}

rmsehist_gpp <- ggplot(fits, aes(x = rmse)) +
  geom_histogram(aes(fill = ..x..), bins = 30) +
  scale_fill_gradient(low = "lightblue", high = "darkred") +
  theme_minimal() +
  labs(fill = "RMSE")

```




```{r, echo = F, message = F, include=F, warning=F}
ER_plot <- comp_df %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = ER.true, color = "True")) +
  geom_point(aes(y = ER.true, color = "True")) +

  # Estimated ER
  geom_ribbon(aes(ymin = ER.est - ER.daily.sd,
                  ymax = ER.est + ER.daily.sd,
                  fill = "Modeled"),
              alpha = 0.2) +
  geom_line(aes(y = ER.est, color = "Modeled")) +
  geom_point(aes(y = ER.est, color = "Modeled")) +

  # Customize colors and labels
  scale_color_manual(name = "ER Type",
                     values = c("True" = "purple4", "Modeled" = "goldenrod")) +
  scale_fill_manual(name = "ER Type",
                    values = c("True" = "purple4", "Modeled" = "goldenrod")) +

  theme_bw() +
  ylab(expression(ER~(g~O[2]~m^-2~d^-1)))

```

```{r, echo = F, message = F, include=T, warning=F}
fits <- comp_df%>% 
  group_by(date)%>%
  summarize(rmse = sqrt(mean((ER.true-ER.est)^2)),
              sd = sd(ER.true),
              min = min(ER.true, na.rm = T),
              max = max(ER.true, na.rm = T),
              range = range(max-min, na.rm = T),
              nrmse = rmse/range)
```

```{r, echo = F, message = F, include=F, warning=F}

rmsehist_ER <- ggplot(fits, aes(x = rmse)) +
  geom_histogram(aes(fill = ..x..), bins = 30) +
  scale_fill_gradient(low = "lightblue", high = "darkred") +
  theme_minimal() +
  labs(fill = "RMSE")

```


```{r, echo = F, message = F, include=F, warning=F}
K600_plot <- comp_df %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = K600.true, color = "True")) +
  geom_point(aes(y = K600.true, color = "True")) +

  # Estimated K600
  geom_ribbon(aes(ymin = K600.est - K600.daily.sd,
                  ymax = K600.est + K600.daily.sd,
                  fill = "Modeled"),
              alpha = 0.2) +
  geom_line(aes(y = K600.est, color = "Modeled")) +
  geom_point(aes(y = K600.est, color = "Modeled")) +

  # Customize colors and labels
  scale_color_manual(name = "K600 Type",
                     values = c("True" = "steelblue3", "Modeled" = "goldenrod")) +
  scale_fill_manual(name = "K600 Type",
                    values = c("True" = "steelblue3", "Modeled" = "goldenrod")) +

  theme_bw() +
  ylab(expression(K600~(g~O[2]~m^-2~d^-1)))

```

```{r, echo = F, message = F, include=T, warning=F}
fits <- comp_df%>% 
  group_by(date)%>%
  summarize(rmse = sqrt(mean((K600.true-K600.est)^2)),
              sd = sd(K600.true),
              min = min(K600.true, na.rm = T),
              max = max(K600.true, na.rm = T),
              range = range(max-min, na.rm = T),
              nrmse = rmse/range)
```

```{r, echo = F, message = F, include=F, warning=F}

rmsehist_K600 <- ggplot(fits, aes(x = rmse)) +
  geom_histogram(aes(fill = ..x..), bins = 30) +
  scale_fill_gradient(low = "lightblue", high = "darkred") +
  theme_minimal() +
  labs(fill = "RMSE")

```

```{r, echo = F, message = F, include=T}
ggarrange(
  GPP_plot,
  rmsehist_gpp, 
  ER_plot,
  rmsehist_ER, 
  K600_plot,
  rmsehist_K600,
  ncol = 2, nrow = 3,
  widths = c(0.7, 0.3),
  common.legend = T,
  align=c("v")
)
```


