---
title: "Streammetabolizer vingette: Simulate dissolved oxygen data from metabolism data"
author: "USGS-R -> KAL"
date: "2025-04-17"

output:
   html_document:
    theme: simplex
    highlight: tango
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type="text/css">
body, td {font-size: 13px;}
code.r{font-size: 9px;}
pre {font-size: 11px}
</style>
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = F, message = F)
knitr::opts_knit$set(root.dir = '/Users/kellyloria/Documents/Publications/')
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```


```{r, echo = F, message = F, include=FALSE}
### Packages
lapply(c("dplyr","ggplot2", "lubridate","tidyverse", "streamMetabolizer", "patchwork"), require, character.only=T)
```


## This builds off the existing vinegette for metabolism data simulation

Available here: https://rdrr.io/github/USGS-R/streamMetabolizer/man/metab_sim.html


### Nothing new here. 
Just creating to help understand streamMetabolizer's built in simulation examples.



### 1. Simulations with variation all at sub-daily scale

Prepare input data (DO used only to pick first DO of each day)

```{r, echo = T, message = F, include=T}
dat <- data_metab('10', res='15')
head(dat)
```

Where site '10' is a built in example and the temporal resolution is 15 muinutes

#### Create true metabolism estimates:

GPP = 3, ER = –4, K600 = 19

```{r, echo = T, message = F, include=T}

dat_daily <- data.frame(date=as.Date(paste0("2012-09-", 18:20)),
                        GPP.daily=3, ER.daily=-4, K600.daily=19, stringsAsFactors=FALSE)

head(dat_daily)
```


#### Set up simulation model

metab_sim {streamMetabolizer}
Takes input data in the form of a sub-daily time series (data) of DO.sat, depth, temperature, and light, and a daily time series (data_daily) of GPP, ER, and K600 values, and turns these into simulated DO.obs. Either data or data_daily should specify a starting DO.obs value for each day; if in data, this takes the form of a DO.obs column with values on at least the first time point of each day (all other values are ignored), or if in data_daily, this takes the form of a DO.mod.1 column with one starting DO value per day.

```{r, echo = T, message = F, include=T}
# define simulation parameters
mm <- metab_sim(
  specs(mm_name('sim'), err_obs_sigma=0.1, err_proc_sigma=2,
        GPP_daily=NULL, ER_daily=NULL, K600_daily=NULL),
  data=dat, data_daily=dat_daily)

get_params(mm)
```

#### Metab estimates are for data without errors:

```{r, echo = T, message = F, include=T}

# Returns the true parameters (GPP, ER, K600) used for simulation
predict_metab(mm) 
```

```{r, echo = T, message = F, include=T}
# Estimates metabolism based on noise-free DO simulation (returns same values as data_daily).
predict_DO(mm)[seq(1,50,by=10),]
```

#### Start simulation: 

Actual simulation called by predict_DO()

```{r, echo = T, message = F, include=T}

mm@specs$sim_seed <- 236
predict_DO(mm)$DO.obs[seq(1,50,by=10)]

# different each time (?) -is it?
predict_DO(mm)$DO.obs[seq(1,50,by=10)]

```


### 2. Simulation with a light-saturation

Defines GPP using the light saturation function:
GPP = Pmax * tanh(alpha * light)

```{r, echo = T, message = F, include=T}

# fancy GPP equation
dat_daily <- data.frame(
  date = as.Date(paste0("2012-09-", 18:20)),
  Pmax = 8,
  alpha = 0.01,
  ER.daily = -4,
  K600.daily = 19,
  stringsAsFactors = FALSE
)

head(dat_daily)
```

Sets up simulation with GPP_fun='satlight' and pulls Pmax/alpha from data_daily.

```{r, echo = T, message = F, include=T}
mm <- metab_sim(
  specs(mm_name('sim', GPP_fun='satlight'), err_obs_sigma=0.1, err_proc_sigma=2,
        Pmax=NULL, alpha=NULL, ER_daily=NULL, K600_daily=NULL),
  data=dat, data_daily=dat_daily)
```

### 3. Simulation with sub-daily and multi-day variability

```{r, echo = T, message = F, include=T}

sp <- specs(mm_name('sim', pool_K600='none'),
            K600_daily = function(n, ...) pmax(0, rnorm(n, 10, 3)))

mm <- metab(sp, dat)
get_params(mm)
predict_metab(mm) # metab estimates are for data without errors
predict_DO(mm)[seq(1,50,by=10),]

```

### 4. Simulation with K~Q relationship

```{r, echo = T, message = F, include=T}
dat <- data_metab('10','15')
sp <- specs(mm_name('sim', pool_K600='binned'))
mm <- metab(sp, dat)
pars <- get_params(mm)
attr(pars, 'K600_eqn')

```
#### Visualize:


Simulated DO (no error) - This shows the ideal DO signal, not influenced by any simulated noise.

Noise in DO.obs - plot includes observation + process errors.


```{r, echo = T, message = F, include=T}
# Plot the underlying "true" DO signal (no error)
p_truth <- plot_DO_preds(mm) + ggtitle("Simulated DO (no error)")

# Plot the modeled predictions vs. noisy DO.obs
p_noisy <- plot_DO_preds(predict_DO(mm)) + ggtitle("Noise in DO.obs")

p_truth + p_noisy

```

